Lossy compression of DNA sequencing quality data

### 2015-04-13 manuscript/q-compression.tex
A few small changes in text (e.g., fraction -> proportion).

Setting all bases as minimum quality results in the highest fraction of mapped reads 88.1\%.
Conversely, setting all bases as maximum quality results in the lowest fraction of mapped reads 72.8\%.

\begin{table*}[!tbhp]
\centering
\caption[]{Quality filtering/trimming for the uncompressed and
  decompressed \textsc{fastq} files. Numbers correspond to the number
  of reads that are kept or discarded after applying compression with
  respect to the original, uncompressed values.}
\label{tab:quality_control}
\begin{tabular}{lr|cc|cc|cc|cc|cc}
 & & \multicolumn{2}{c|}{{\em MaxQual}} & \multicolumn{2}{c|}{{\em MinQual}} & \multicolumn{2}{c|}{{\em 2-bin}} & \multicolumn{2}{c|}{{\em Degree 0}} & \multicolumn{2}{c}{{\em Degree 1}} \\
 & & kept & discarded & kept & discarded & kept & discarded & kept & discarded & kept & discarded \\ 
  \cline{2-12}
& kept & 883576 &   0 &   0 & 883576 & 871035 & 12541 & 728349 & 155227 & 871510 & 12066 \\ 
{\em original} & discarded & 141858 &   0 &   0 & 141858 & 1882 & 139976 & 1475 & 140383 & 9214 & 132644 \\ 
\cline{2-12}
 & fraction & 1.000 & 0.000 & 0.000 & 1.000 & 0.851 & 0.149 & 0.712 & 0.288 & 0.859 & 0.141 \\ 
\end{tabular}

@@

\begin{tabular}{lr|cc|cc|cc|cc|cc}
 & & \multicolumn{2}{c|}{{\em Degree 3}} & \multicolumn{2}{c|}{{\em Degree 5}} & \multicolumn{2}{c|}{{\em Degree 7}} & \multicolumn{2}{c|}{{\em Profile (64)}} & \multicolumn{2}{c}{{\em Profile (128)}} \\
 & & kept & discarded & kept & discarded & kept & discarded & kept & discarded & kept & discarded \\ 
\cline{2-12}
& kept & 872006 & 11570 & 878936 & 4640 & 877802 & 5774 & 876149 & 7427 & 877513 & 6063 \\ 
{\em original}  &  discarded & 10369 & 131489 & 10004 & 131854 & 8821 & 133037 & 15795 & 126063 & 17453 & 124405 \\
\cline{2-12}
& fraction & 0.860 & 0.140 & 0.867 & 0.133 & 0.865 & 0.135 & 0.870 & 0.130 & 0.873 & 0.127 \\ 
\end{tabular}

@@

\begin{tabular}{lr|cc|cc|cc|cc|cc}
&  & \multicolumn{2}{c|}{{\em Profile (256)}} & \multicolumn{2}{c|}{{\em QualComp (6)}} & \multicolumn{2}{c|}{{\em QualComp (10)}} & \multicolumn{2}{c|}{{\em QualComp (30)}} & \multicolumn{2}{c}{{\em QualComp (100)}} \\
 & &  kept & discarded & kept & discarded & kept & discarded & kept & discarded & kept & discarded \\ 
\cline{2-12}
& kept & 881215 & 2361 & 878702 & 4874 & 879682 & 3894 & 881948 & 1628 & 881939 & 1637 \\ 
{\em original}  & discarded & 15985 & 125873 & 13434 & 128424 & 12855 & 129003 & 10872 & 130986 & 7179 & 134679 \\ 
\cline{2-12}
& fraction & 0.875 & 0.125 & 0.870 & 0.130 & 0.870 & 0.130 & 0.871 & 0.129 & 0.867 & 0.133 \\ 
\end{tabular}
\end{table*}

\begin{table*}[!tbhp]
\centering
\caption[]{Mapping results of decompressed \textsc{fastq} files
  against \textit{Rhodobacter sphaeroides} reference genome using
  Bowtie2. Numbers corresponds to the proportion of mapped reads with
  respect to the uncompressed \textsc{fastq}. ``Shared'' denotes the
  percentage of mapped reads by both the uncompressed and decompressed
  data. ``Uncompressed only'' denotes the percentage of reads mapped
  from the uncompressed data that are not mapped after
  decompression. ``Compressed only'' denotes the percentage of reads
  mapped from the decompressed data that were not mapped before
  compression.}
\begin{tabular}{lr|cc|cc|cc|cc|cc}
 & & \multicolumn{2}{c|}{{\em MaxQual}} & \multicolumn{2}{c|}{{\em MinQual}} & \multicolumn{2}{c|}{{\em 2-bin}} & \multicolumn{2}{c|}{{\em Degree 0}} & \multicolumn{2}{c}{{\em Degree 1}} \\
& & mapped & unmapped & mapped & unmapped & mapped & unmapped & mapped & unmapped & mapped & unmapped \\ 
\cline{2-12}
& mapped & 746716 & 145897 & 892613 &   0 & 891864 & 749 & 851682 & 40931 & 883390 & 9223 \\ 
{\em original} & unmapped &   0 & 132821 & 10821 & 122000 & 186 & 132635 &  67 & 132754 &  55 & 132766 \\ 
\cline{2-12}
%& sum & 746716 & 278718 & 903434 & 122000 & 892050 & 133384 & 851749 & 173685 & 883445 & 141989 \\ 
& fraction & 0.728 & 0.272 & 0.881 & 0.119 & 0.870 & 0.130 & 0.831 & 0.169 & 0.862 & 0.138 \\ 
\end{tabular}

@@

\begin{tabular}{lr|cc|cc|cc|cc|cc}
 & & \multicolumn{2}{c|}{{\em Degree 3}} & \multicolumn{2}{c|}{{\em Degree 5}} & \multicolumn{2}{c|}{{\em Degree 7}} & \multicolumn{2}{c|}{{\em Profile (64)}} & \multicolumn{2}{c}{{\em Profile (128)}} \\
& & mapped & unmapped & mapped & unmapped & mapped & unmapped & mapped & unmapped & mapped & unmapped \\ 
\cline{2-12}
& mapped & 889537 & 3076 & 891019 & 1594 & 891479 & 1134 & 891753 & 860 & 891952 & 661 \\ 
{\em original} & unmapped & 117 & 132704 & 155 & 132666 & 154 & 132667 & 144 & 132677 & 143 & 132678 \\ 
\cline{2-12}
%& sum & 889654 & 135780 & 891174 & 134260 & 891633 & 133801 & 891897 & 133537 & 892095 & 133339 \\ 
& fraction & 0.868 & 0.132 & 0.869 & 0.131 & 0.870 & 0.130 & 0.870 & 0.130 & 0.870 & 0.130 \\ 
\end{tabular}

@@

\begin{tabular}{lr|cc|cc|cc|cc|cc}
&  & \multicolumn{2}{c|}{{\em Profile (256)}} & \multicolumn{2}{c|}{{\em QualComp (6)}} & \multicolumn{2}{c|}{{\em QualComp (10)}} & \multicolumn{2}{c|}{{\em QualComp (30)}} & \multicolumn{2}{c}{{\em QualComp (100)}} \\
& &  mapped & unmapped & mapped & unmapped & mapped & unmapped & mapped & unmapped & mapped & unmapped \\ 
\cline{2-12}
& mapped & 892051 & 562 & 891375 & 1238 & 891777 & 836 & 892233 & 380 & 892454 & 159 \\ 
{\em original}  & unmapped & 119 & 132702 & 304 & 132517 & 265 & 132556 & 220 & 132601 & 172 & 132649 \\ 
\cline{2-12}
%& sum & 892170 & 133264 & 891679 & 133755 & 892042 & 133392 & 892453 & 132981 & 892626 & 132808 \\ 
& fraction & 0.870 & 0.130 & 0.870 & 0.130 & 0.870 & 0.130 & 0.870 & 0.130 & 0.870 & 0.130 \\

>>>>

Setting all bases as minimum quality results in the highest proportion of mapped reads 88.1\%.
Conversely, setting all bases as maximum quality results in the lowest proportion of mapped reads 72.8\%.

\begin{table*}[!tbhp]
\centering
\caption[]{Quality filtering/trimming for the uncompressed and
  decompressed \textsc{fastq} files. Numbers correspond to the number
  of reads that are kept or discarded after applying compression with
  respect to the original, uncompressed values.}
\label{tab:quality_control}
\begin{tabular}{lr|cc|cc|cc|cc|cc}
 & & \multicolumn{2}{c|}{{\em MaxQual}} & \multicolumn{2}{c|}{{\em MinQual}} & \multicolumn{2}{c|}{{\em 2-bin}} & \multicolumn{2}{c|}{{\em Degree 0}} & \multicolumn{2}{c}{{\em Degree 1}} \\
 & & kept & discarded & kept & discarded & kept & discarded & kept & discarded & kept & discarded \\ 
  \cline{2-12}
& kept & 883576 &   0 &   0 & 883576 & 871035 & 12541 & 728349 & 155227 & 871510 & 12066 \\ 
{\em original} & discarded & 141858 &   0 &   0 & 141858 & 1882 & 139976 & 1475 & 140383 & 9214 & 132644 \\ 
\cline{2-12}
 & proportion & 1.000 & 0.000 & 0.000 & 1.000 & 0.851 & 0.149 & 0.712 & 0.288 & 0.859 & 0.141 \\ 
\end{tabular}

@@

\begin{tabular}{lr|cc|cc|cc|cc|cc}
 & & \multicolumn{2}{c|}{{\em Degree 3}} & \multicolumn{2}{c|}{{\em Degree 5}} & \multicolumn{2}{c|}{{\em Degree 7}} & \multicolumn{2}{c|}{{\em Profile (64)}} & \multicolumn{2}{c}{{\em Profile (128)}} \\
 & & kept & discarded & kept & discarded & kept & discarded & kept & discarded & kept & discarded \\ 
\cline{2-12}
& kept & 872006 & 11570 & 878936 & 4640 & 877802 & 5774 & 876149 & 7427 & 877513 & 6063 \\ 
{\em original}  &  discarded & 10369 & 131489 & 10004 & 131854 & 8821 & 133037 & 15795 & 126063 & 17453 & 124405 \\
\cline{2-12}
& proportion & 0.860 & 0.140 & 0.867 & 0.133 & 0.865 & 0.135 & 0.870 & 0.130 & 0.873 & 0.127 \\ 
\end{tabular}

@@

\begin{tabular}{lr|cc|cc|cc|cc|cc}
&  & \multicolumn{2}{c|}{{\em Profile (256)}} & \multicolumn{2}{c|}{{\em QualComp (6)}} & \multicolumn{2}{c|}{{\em QualComp (10)}} & \multicolumn{2}{c|}{{\em QualComp (30)}} & \multicolumn{2}{c}{{\em QualComp (100)}} \\
 & &  kept & discarded & kept & discarded & kept & discarded & kept & discarded & kept & discarded \\ 
\cline{2-12}
& kept & 881215 & 2361 & 878702 & 4874 & 879682 & 3894 & 881948 & 1628 & 881939 & 1637 \\ 
{\em original}  & discarded & 15985 & 125873 & 13434 & 128424 & 12855 & 129003 & 10872 & 130986 & 7179 & 134679 \\ 
\cline{2-12}
& proportion & 0.875 & 0.125 & 0.870 & 0.130 & 0.870 & 0.130 & 0.871 & 0.129 & 0.867 & 0.133 \\ 
\end{tabular}
\end{table*}

\begin{table*}[!tbhp]
\centering
\caption[]{Mapping results of decompressed \textsc{fastq} files
  against \textit{Rhodobacter sphaeroides} reference genome using
  Bowtie2. Numbers corresponds to the proportion of mapped reads with
  respect to the uncompressed \textsc{fastq}. ``Shared'' denotes the
  percentage of mapped reads by both the uncompressed and decompressed
  data. ``Uncompressed only'' denotes the percentage of reads mapped
  from the uncompressed data that are not mapped after
  decompression. ``Compressed only'' denotes the percentage of reads
  mapped from the decompressed data that were not mapped before
  compression.}
\begin{tabular}{lr|cc|cc|cc|cc|cc}
 & & \multicolumn{2}{c|}{{\em MaxQual}} & \multicolumn{2}{c|}{{\em MinQual}} & \multicolumn{2}{c|}{{\em 2-bin}} & \multicolumn{2}{c|}{{\em Degree 0}} & \multicolumn{2}{c}{{\em Degree 1}} \\
& & mapped & unmapped & mapped & unmapped & mapped & unmapped & mapped & unmapped & mapped & unmapped \\ 
\cline{2-12}
& mapped & 746716 & 145897 & 892613 &   0 & 891864 & 749 & 851682 & 40931 & 883390 & 9223 \\ 
{\em original} & unmapped &   0 & 132821 & 10821 & 122000 & 186 & 132635 &  67 & 132754 &  55 & 132766 \\ 
\cline{2-12}
%& sum & 746716 & 278718 & 903434 & 122000 & 892050 & 133384 & 851749 & 173685 & 883445 & 141989 \\ 
& proportion & 0.728 & 0.272 & 0.881 & 0.119 & 0.870 & 0.130 & 0.831 & 0.169 & 0.862 & 0.138 \\ 
\end{tabular}

@@

\begin{tabular}{lr|cc|cc|cc|cc|cc}
 & & \multicolumn{2}{c|}{{\em Degree 3}} & \multicolumn{2}{c|}{{\em Degree 5}} & \multicolumn{2}{c|}{{\em Degree 7}} & \multicolumn{2}{c|}{{\em Profile (64)}} & \multicolumn{2}{c}{{\em Profile (128)}} \\
& & mapped & unmapped & mapped & unmapped & mapped & unmapped & mapped & unmapped & mapped & unmapped \\ 
\cline{2-12}
& mapped & 889537 & 3076 & 891019 & 1594 & 891479 & 1134 & 891753 & 860 & 891952 & 661 \\ 
{\em original} & unmapped & 117 & 132704 & 155 & 132666 & 154 & 132667 & 144 & 132677 & 143 & 132678 \\ 
\cline{2-12}
%& sum & 889654 & 135780 & 891174 & 134260 & 891633 & 133801 & 891897 & 133537 & 892095 & 133339 \\ 
& proportion & 0.868 & 0.132 & 0.869 & 0.131 & 0.870 & 0.130 & 0.870 & 0.130 & 0.870 & 0.130 \\ 
\end{tabular}

@@

\begin{tabular}{lr|cc|cc|cc|cc|cc}
&  & \multicolumn{2}{c|}{{\em Profile (256)}} & \multicolumn{2}{c|}{{\em QualComp (6)}} & \multicolumn{2}{c|}{{\em QualComp (10)}} & \multicolumn{2}{c|}{{\em QualComp (30)}} & \multicolumn{2}{c}{{\em QualComp (100)}} \\
& &  mapped & unmapped & mapped & unmapped & mapped & unmapped & mapped & unmapped & mapped & unmapped \\ 
\cline{2-12}
& mapped & 892051 & 562 & 891375 & 1238 & 891777 & 836 & 892233 & 380 & 892454 & 159 \\ 
{\em original}  & unmapped & 119 & 132702 & 304 & 132517 & 265 & 132556 & 220 & 132601 & 172 & 132649 \\ 
\cline{2-12}
%& sum & 892170 & 133264 & 891679 & 133755 & 892042 & 133392 & 892453 & 132981 & 892626 & 132808 \\ 
& proportion & 0.870 & 0.130 & 0.870 & 0.130 & 0.870 & 0.130 & 0.870 & 0.130 & 0.870 & 0.130 \\
